Exploring the tree of numerical semigroups

### 2013-10-07 Papier/article.tex
Rereading + more computed values

$iii)$ $c(S)=1+\max(\NN\setminus S)$, the \emph{conductor} of $S$ for $S$ different from $\NN$. By convention the conductor of $\NN$ is $0$. 
\end{defi}

@@

We construct the tree of numerical semigroups, denoted by $\mathcal{T}$ as follows. 
The root of the tree is the unique semigroup of genus $0$, \ie, $\left<1\right>$ that is equal to $\NN$. 
If~$S$ is a semigroup in the tree,  the sons of $S$ are exactly the semigroup $S^x$ where $x$ belongs to $\Irr(S)\cap[c(s),+\infty[$. 
By convention, when depicting the tree, the ,numerical semigroup $S^x$ is in the left of $S^y$ if $x$ is smaller than $y$.

@@

As $\NN\setminus{S}$ contains $\{1,..,m(S)-1\}$ we have 
\begin{equation}
\label{E:m}
m(S)\leq g(S)+1.
\end{equation}
Let $x$ be an element of $S\cap[0,c(S)-1]$. 
The integer $y=c(S)-1-x$ lies in $[0,c(S)-1]$. 
Moreover $y$ is not in $S$, for otherwise  we write $c(S)-1=y+x$ where $x,y$ are elements of $S$ implying $c(S)-1\in S$. 
In contradiction with the definition of conductor. 
Thus we define an involution 
\[
\begin{array}{rcl}
\psi:[0,c(S)-1]&\to&[0,c(S)-1]\\
x&\mapsto&c(S)-1-x
\end{array}
\]
that send $S\cap[0,c(S)-1]$ into $[0,c(S)-1]\setminus S$.
The cardinality of $[0,c(S)[\setminus S$ is exactly $g(S)$. Let us denote by $k$ the cardinality  of $S\cap[0,c(S)-1]$. 
Since $\psi$ is injective we must have $k\leq g(S)$. 
>From the relation 
\[
[0,c(S)[=S\cap[0,c(S)[\sqcup[0,c(S)[\setminus S
\]
 we obtain $c(S)=k+g(S)$ and so 
\begin{equation}
\label{E:c}
c(S)\leq 2 g(S).
\end{equation}

@@

\section{A new algorithm}
\label{S:Algo}
We can easily explore the tree of numerical semigroups up to a genus $G$ using a depth search first algorithm using a stack (see~\ref{A:Exp}). 
This approach does not seem to have been used before. 
In particular,  M. Bras-Amor\'os and M. Delgado use instead a breadth search first exploration. 
The main advantage in our approach is the small memory needs. 
The cost to pay is that, if we want to explore the tree deepero, we must restart from the root.

@@

\begin{minipage}{0.5\textwidth}
\begin{tabular}{|r|r|r|}
\hline
g & $n_g$ & $n_g/n_{g-1}$\\
\hline
0 & 1 &  \\
1 & 1 & 1.0 \\
2 & 2 & 2.0 \\
3 & 4 & 2.0 \\
4 & 7 & 1.75 \\
5 & 12 & 1.71428 \\
6 & 23 & 1.91666 \\
7 & 39 & 1.69565 \\
8 & 67 & 1.71794 \\
9 & 118 & 1.76119 \\
10 & 204 & 1.72881 \\
11 & 343 & 1.68137 \\
12 & 592 & 1.72594 \\
13 & 1001 & 1.69087 \\
14 & 1693 & 1.69130 \\
15 & 2857 & 1.68753 \\
16 & 4806 & 1.68218 \\
17 & 8045 & 1.67394 \\
18 & 13467 & 1.67395 \\
19 & 22464 & 1.66807 \\
20 & 37396 & 1.66470 \\
21 & 62194 & 1.66311 \\
22 & 103246 & 1.66006 \\
23 & 170963 & 1.65588 \\
24 & 282828 & 1.65432 \\
25 & 467224 & 1.65197 \\
26 & 770832 & 1.64981 \\
27 & 1270267 & 1.64791 \\
28 & 2091030 & 1.64613 \\
29 & 3437839 & 1.64408 \\
\hline
\end{tabular}
\end{minipage}
\begin{minipage}{0.5\textwidth}
\begin{tabular}{|r|r|r|}
\hline
g & $n_g$ & $n_g/n_{g-1}$\\
\hline
30 & 5646773 & 1.64253 \\
31 & 9266788 & 1.64107 \\
32 & 15195070 & 1.63973 \\
33 & 24896206 & 1.63843 \\
34 & 40761087 & 1.63724 \\
35 & 66687201 & 1.63605 \\
36 & 109032500 & 1.63498 \\
37 & 178158289 & 1.63399 \\
38 & 290939807 & 1.63304 \\
39 & 474851445 & 1.63212 \\
40 & 774614284 & 1.63127 \\
41 & 1262992840 & 1.63047 \\
42 & 2058356522 & 1.62974 \\
43 & 3353191846 & 1.62906 \\
44 & 5460401576 & 1.62841 \\
45 & 8888486816 & 1.62780 \\
46 & 14463633648 & 1.62723 \\
47 & 23527845502 & 1.62668 \\
48 & 38260496374 & 1.62617 \\
49 & 62200036752 & 1.62569 \\
50 & 101090300128 & 1.62524 \\
51 & 164253200784 & 1.62481 \\
52 & 266815155103 & 1.62441 \\
53 & 433317458741 & 1.62403 \\
54 & 703569992121 & 1.62368 \\
55 & 1142140736859 & 1.62335 \\
56 & 1853737832107 & 1.62303 \\
57 & 3008140981820 & 1.62274 \\
58 & 4880606790010 & 1.62246 \\
59 & 7917344087695 & 1.62220 \\
60 & 12841603251351 & 1.62195 \\
\hline
\end{tabular}
\end{minipage}
\vspace{2em}

>>>>

$iii)$ $c(S)=1+\max(\NN\setminus S)$, the \emph{conductor} of $S$ for $S$ different from $\NN$. By convention, the conductor of $\NN$ is $0$. 
\end{defi}

@@

We construct the tree of numerical semigroups, denoted by $\mathcal{T}$ as follows. 
The root of the tree is the unique semigroup of genus $0$, \ie, $\left<1\right>$ that is equal to $\NN$. 
If~$S$ is a semigroup in the tree,  the sons of $S$ are exactly the semigroup $S^x$ where $x$ belongs to $\Irr(S)\cap[c(s),+\infty[$. 
By convention, when depicting the tree, the numerical semigroup $S^x$ is in the left of $S^y$ if $x$ is smaller than $y$.

@@

As $\NN\setminus{S}$ contains $\{1,..,m(S)-1\}$ we have 
\begin{equation}
\label{E:m}
m(S)\leq g(S)+1.
\end{equation}
Let $x$ be an element of $S\cap[0,c(S)-1]$. 
The integer $y=c(S)-1-x$ lies in $[0,c(S)-1]$. 
Moreover $y$ is not in $S$, for otherwise  we write $c(S)-1=y+x$ where $x,y$ are elements of $S$ implying $c(S)-1\in S$. 
In contradiction with the definition of conductor. 
Thus we define an involution 
\[
\begin{array}{rcl}
\psi:[0,c(S)-1]&\to&[0,c(S)-1]\\
x&\mapsto&c(S)-1-x
\end{array}
\]
that send $S\cap[0,c(S)-1]$ into $[0,c(S)-1]\setminus S$.
The cardinality of $[0,c(S)[\setminus S$ is exactly $g(S)$. Let us denote by $k$ the cardinality  of $S\cap[0,c(S)-1]$. 
Since $\psi$ is injective we must have $k\leq g(S)$. 
From the relation 
\[
[0,c(S)[=S\cap[0,c(S)[\sqcup[0,c(S)[\setminus S
\]
 we obtain $c(S)=k+g(S)$ and so 
\begin{equation}
\label{E:c}
c(S)\leq 2 g(S).
\end{equation}

@@

\section{A new algorithm}
\label{S:Algo}
We can easily explore the tree of numerical semigroups up to a genus $G$ using a depth search first algorithm using a stack (see~\ref{A:Exp}). 
This approach does not seem to have been used before. 
In particular,  M. Bras-Amor\'os and M. Delgado use instead a breadth search first exploration. 
The main advantage in our approach is the small memory needs. 
The cost to pay is that, if we want to explore the tree deeper, we must restart from the root.

@@

\begin{minipage}{0.5\textwidth}
\begin{tabular}{|r|r|r|}
\hline
g & $n_g$ & $n_g/n_{g-1}$\\
\hline
0 & 1 &  \\
1 & 1 & 1.0 \\
2 & 2 & 2.0 \\
3 & 4 & 2.0 \\
4 & 7 & 1.75 \\
5 & 12 & 1.71428 \\
6 & 23 & 1.91666 \\
7 & 39 & 1.69565 \\
8 & 67 & 1.71794 \\
9 & 118 & 1.76119 \\
10 & 204 & 1.72881 \\
11 & 343 & 1.68137 \\
12 & 592 & 1.72594 \\
13 & 1001 & 1.69087 \\
14 & 1693 & 1.69130 \\
15 & 2857 & 1.68753 \\
16 & 4806 & 1.68218 \\
17 & 8045 & 1.67394 \\
18 & 13467 & 1.67395 \\
19 & 22464 & 1.66807 \\
20 & 37396 & 1.66470 \\
21 & 62194 & 1.66311 \\
22 & 103246 & 1.66006 \\
23 & 170963 & 1.65588 \\
24 & 282828 & 1.65432 \\
25 & 467224 & 1.65197 \\
26 & 770832 & 1.64981 \\
27 & 1270267 & 1.64791 \\
28 & 2091030 & 1.64613 \\
29 & 3437839 & 1.64408 \\
30 & 5646773 & 1.64253 \\
31 & 9266788 & 1.64107 \\
32 & 15195070 & 1.63973 \\
33 & 24896206 & 1.63843 \\
34 & 40761087 & 1.63724 \\
35 & 66687201 & 1.63605 \\
\hline
\end{tabular}
\end{minipage}
\begin{minipage}{0.5\textwidth}
\begin{tabular}{|r|r|r|}
\hline
g & $n_g$ & $n_g/n_{g-1}$\\
\hline
36 & 109032500 & 1.63498 \\
37 & 178158289 & 1.63399 \\
38 & 290939807 & 1.63304 \\
39 & 474851445 & 1.63212 \\
40 & 774614284 & 1.63127 \\
41 & 1262992840 & 1.63047 \\
42 & 2058356522 & 1.62974 \\
43 & 3353191846 & 1.62906 \\
44 & 5460401576 & 1.62841 \\
45 & 8888486816 & 1.62780 \\
46 & 14463633648 & 1.62723 \\
47 & 23527845502 & 1.62668 \\
48 & 38260496374 & 1.62617 \\
49 & 62200036752 & 1.62569 \\
50 & 101090300128 & 1.62524 \\
51 & 164253200784 & 1.62481 \\
52 & 266815155103 & 1.62441 \\
53 & 433317458741 & 1.62403 \\
54 & 703569992121 & 1.62368 \\
55 & 1142140736859 & 1.62335 \\
56 & 1853737832107 & 1.62303 \\
57 & 3008140981820 & 1.62274 \\
58 & 4880606790010 & 1.62246 \\
59 & 7917344087695 & 1.62220 \\
60 & 12841603251351 & 1.62195 \\
61 & 20825558002053 & 1.62173 \\
62 & 33768763536686 & 1.62151 \\
63 & 54749244915730 & 1.62130 \\
64 & 88754191073328 & 1.62110 \\
65 & 143863484925550 & 1.62092 \\
66 & 233166577125714 & 1.62075 \\
67 & 377866907506273 & 1.62059 \\\hline
\end{tabular}
\end{minipage}
\vspace{2em}
